### 2011-12-16 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@

@@

\STATE $type = null$
\IF{$exprNode$ represents a literal value}
     \STATE $type =$ particular literal value type, exactly-once cardinality
\ELSIF{$exprNode$ represents a call of function $f$}
     \STATE $type = getFunctionReturnType(f)$
\ELSIF{$exprNode$ represents variable reference $var$}
     \STATE $type = getVariableType(contextVarTypes, var)$
\ELSIF{$exprNode$ represents path expression $path$}
     \IF{$path$ selects an element set}
         \STATE $type =$ element type, zero-or-more cardianlity, $path$ included
     \ELSIF{$path$ selects an attribute set}
         \STATE $type =$ attribute type, zero-or-more cardianlity, $path$ included
     \ELSE
         \STATE $type = null$
     \ENDIF
\ELSIF{$exprNode$ represents operator $op$}
     \IF{$op$ is a comparison operator}
         \STATE $type = xs:boolean$
     \ELSIF{$op$ is an additive operator $AND$ one of the operands is of a numeric type}
         \IF{both operands are of numeric types}
             \STATE $type =$ the numberic type which is more or equally general than the other
         \ELSE
             \STATE $type =$ type of the numeric operand
         \ENDIF
     \ENDIF
\ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
     \STATE  $type =$ type of an expression in the return clause of $flwor$
\ENDIF

@@

\todo[inline]{algoritmus 6.2, riadok 2, slovo "represents" - Tady by neškodilo definovat nějakou funkci a její možné návratové hodnoty a toto používat.}
\todo[inline]{algoritmus 6.2, riadok 19 - Totéž, neškodila by nějaká funkce, tj. trochu formálnější zápis. Jednak se pravděpodobně bude hodit dál a druhak to zpřehlední popis.}

>>>>

\todo[inline]{V ExprNode si pamatame aj typ vyrazu}

@@

Some pseudo code algorithms in the following sections need to determine the type of a node instance. For this purpose, we define function $isA(nodeInstance, nodeType)$. The first argument is an instance of any node and the second is a node type itself. The function returns boolean value $true$ if the instance is of the specified node type, according to the described principle. Otherwise, it returns $false$.

For instance, assuming variable $literal$ is an instance of node $LiteralNode$, calls $isA(literal, LiteralNode)$, $isA(literal, ExprNode)$ return $true$, while call $isA(literal, ContentNode)$ returns $false$.

@@

\STATE $type = null$
\IF{$isA(exprNode, LiteralNode)$}
     \STATE $type =$ ((LiteralNode)exprNode).type, exactly-once cardinality
\ELSIF{$isA(exprNode, FunctionCallNode)$}
     \STATE $type = getFunctionReturnType(((FunctionCallNode)exprNode).functionName)$
\ELSIF{$isA(exprNode, VarRefNode)$}
     \STATE $type = getVariableType(contextVarTypes, ((VarRefNode)exprNode).variableName)$
\ELSIF{$isA(exprNode, PathExprNode)$}
     \STATE $path = ((PathExprNode)exprNode).path$
     \IF{$path$ selects an element set}
         \STATE $type =$ element type, zero-or-more cardianlity, $path$ included
     \ELSIF{$path$ selects an attribute set}
         \STATE $type =$ attribute type, zero-or-more cardianlity, $path$ included
     \ELSE
         \STATE $type = null$
     \ENDIF
\ELSIF{$isA(exprNode, OperatorNode)$}
     \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
\ELSIF{$isA(exprNode, FLWORExprNode)$}
     \STATE $returnClauseNode = ((FLWORExprNode)exprNode).returnClauseNode$
     \STATE $type = returnClauseNode.type$
\ENDIF

@@

\todo[inline]{Napisat, ze determineOperatorType() tiez nie je kompletna}
\todo[inline]{definovat isOperatorClassAddition()}
\todo[inline]{definovat isNumericType()}
\todo[inline]{definovat isOperatorClassComparison()}

\begin{algorithm}
\caption{Function determineOperatorType}
\label{ALG_func_determineOperatorType}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $operatorNode$: An operator expression node in the syntax tree. \\
    $contextVarTypes$: Types of local variables valid in the current subtree.
}
\ENSURE The operator expression type.

\STATE $type = null$
\STATE $operator = operatorNode.operator$
\IF{$isOperatorClassComparison(operator)$}
    \STATE $type = xs:boolean$
\ELSIF{$isOperatorClassAddition(operator) AND$ one of the operands is of a numeric type}
    \STATE $leftOperand = ((OperatorNode)exprNode).leftOperand$
    \STATE $rightOperand = ((OperatorNode).exprNode).rightOperand$
    \IF{$isNumericType(leftOperand.type) AND isNumericType(rightOperand.type)$}
        \STATE $type = selectMoreGeneralType(leftOperand.type, rightOperand.type)$
    \ELSIF{$isNumericType(leftOperand.type)$}
        \STATE $type = leftOperand.type$
    \ELSIF{$isNumericType(rightOperand.type)$}
        \STATE $type = rightOperand.type$
    \ENDIF
\ENDIF
